---
title: 'Appendix 1: Lower Bound of Effective Number of Functions'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The minimum number of effective functions will occur at the minimum level of evenness (maximum level of dominance) for any set of functions. Conveniently, using a standardized scale for functions between 0 and 1, the maximum level any function can achieve is 1. If all functions, save one, are 0, then by definition the number of effective functions will be 1. After this, one function will be 1 and all functions save one other will be 0. As we increase $A$, this function will increase in value until it, too, is 1. At which point, as $A$ increases, a third function will begin increasing in value, etc.

Noting that $AK$, the average level of function multiplied by the total number of functions, is always just greater than the number of functions that are equal to 1, we can a quantity $G$ that is the number of functions equal to 1, and is equivalent to the floor of $AK$ (i.e., $G  = \lfloor AK \rfloor$).

For any level of A, the value of the function whose value is greater than 0 but less than 1 is $AK - G$, as $AK$ is the number of functions that are 1 + the additional contribution of our non-zero other function. Therefore,

$$
\tag{eqn. 1}
\begin{equation}
\begin{aligned} 
&F_{1}...F_{K-G-1} = 0 \\
&F_{K-G} = AK - G \\
&F_{K-G+1}...F_{K} = 1
\end{aligned} 
\end{equation}
$$
*RBnote: I could work out that this worked, but it took me a bit of time. Why not do $F_1 ... F_{G} = 1$, $F_{G + 1} = AK - G$ and $F_{G+2} ... F_{K} = 0$. The formula would still work as $\sum_i F_i = G + AK + 0 = AK  $ *

and therefore summing up all of the functions at the maximum dominance for a given level of $A$,

$$
\tag{eqn. 2}
\sum F_i = 0 + AK - G + G  \\
\Rightarrow AK 
$$
We can use this to determine the lower bound of the effective number of functions. Looking at $p_i$ at maximum dominance


$$
\tag{eqn. 3}
p_i = \frac{F_i}{\sum F_i} \\
\Rightarrow \frac{F_i}{AK} 
$$

We can then substitute this into equation 3.

$$
\tag{eqn. 4}
^{q}N = \left[ \sum_{i=1}^{K} \left(\frac{F_{i}}{AK}\right)^q \right] ^ {1/(1-q)} \\
$$
*RB note: I've changed S to K to be consistent with above*

$$
\tag{eqn. 5}
\Rightarrow \left[{\frac{1}{A^q \cdot K^{q}}} \cdot \sum_{i=1}^{K}F_{i}^q \right]^{1/(1-q)}
$$

At maximum dominance we know that 

$$
\tag{eqn. 6}
\sum_{i=1}^{K}F_{i}^q  = G + (AK-G)^q
$$

So that the lower bound of effective number of functions is

$$
\tag{eqn. 7}
^qN_{lower}(A) = \left[ \frac{G + (AK-G)^q}{A^qK^q} \right]^{1/(1-q)}
$$
*RBcomment: I used Mathematica to derive qN_lower when q -> 1*
For q=1, we can begin by re-arranging the limit by substituting in for $p_i$.

$$
\tag{eqn. 8}
\begin{equation}
\begin{aligned}
&^{1}N_{lower} = \exp\left(-\sum p_i \;\log\; p_i \right) \\
&\Rightarrow \exp\left(-\sum \frac{F_i}{AK} \;\log\; \frac{F_i}{AK} \right)\\
&\Rightarrow \exp\left(-\frac{1}{AK}  \sum (F_i \; \log \; F_i \; - F_i\; \log \; AK) \right) \\
\end{aligned}
\end{equation}
$$
Taking 0 log 0 = 0 and noting that log 1 = 0, we can use our previous exploration of the three possible values for $F_i$ at maximum dominance ($AK - G$) and plug those in to the above equation to get

$$
\begin{equation}
\tag{eqn. 9}
\begin{aligned}
&^{1}N_{lower} =  \exp\left(-\frac{1}{AK}  \Bigl[ (AK-G)\;\log(AK-G) - G \; \log AK - (AK-G) \;\log\; AK \Bigr] \right)\\
&\Rightarrow  \exp\left(\frac{G}{AK}\;\log AK + \frac{AK-G}{AK}\;\log AK - \frac{AK-G}{AK}\; \log(AK-G) \right) \\
&\Rightarrow \frac{-AK(AK-G)^\frac{(AK-G)}{AK}}{(G - AK)}
\end{aligned}
\end{equation}
$$

*I think this can be simplified to*

$$
\begin{equation}
\begin{aligned}
&-AK((AK -G)^(\frac{AK - G}{AK}-1) \\
&\Rightarrow -AK(AK -G)^\frac{-G}{AK}

\end{aligned}
\end{equation}
$$


When AS-G = 0, this reduces to $AK ^ {\frac{G}{AK}}$, rather than having to divide by 0.

*RBcomment: I think this might be wrong. Mathematica helped me derive the expression directly. It is below*
$$
^{1}N_{lower} = A K (A K - G)^{(\frac{G}{AK} - 1)}
$$


We can write a function in R to explore this relationship between $^qN$ and A.

```{r, echo = TRUE}
get_mfn_floor <- function(a, k, q){
  ak <- a*k
  g <- floor(a*k)
  
  if(q==1){
    if(ak-g==0){
      #AK ^ {\frac{G}{AK}}
      return(ak^(g/ak))
    }
    ret <- -1*ak*(ak-g)^(g/ak)/(g-ak)
    ix <- is.nan(ret)
    ret[ix] <- ak[ix]^(g[ix]/ak[ix])#exp(-1/as[ix]*(g[ix]*log(1/as[ix])))
#    return(ret)
  }else{
  #\left[ \frac{G + (AK-G)^q}{A^qK^q} \right]^{1/(1-q)}
  ret <- ( (g+(ak-g)^q)/(a^q * k^q) ) ^ (1/(1-q))
  }
  return(ret)
}
```


We can use this function to then explore the shape of the relationship between A and possible values of $^qN$ which is reported in the text as figure 1.



```{r show_conceptual, message=FALSE, warning=FALSE}

library(dplyr)
library(tidyr)
library(ggplot2)
library(forcats)

n_func <- 4

n_range <- crossing(data.frame(a = seq(0,1,0.01)), 
                    data.frame(q = 1:4)) %>%
  group_by(q, a) %>%
  mutate(n_upper=n_func, n_lower=get_mfn_floor(a, n_func, q))

ggplot(n_range, aes(x=a, y=n_lower, 
                          ymin = n_lower,
                          ymax = n_upper,
                          fill = fct_rev(factor(q)),
                          color= fct_rev(factor(q)))) +
  geom_ribbon(alpha=0.7) +
  xlab("Average Level of Functioning") +
  ylab("Effective Number of Functions") +
  ylim(c(0,n_func)) +
  scale_fill_brewer(palette="Accent") +
  scale_color_brewer(palette="Accent") +
  guides(color="none", fill=guide_legend("Order (q)")) +
  theme_bw(base_size = 14)
  
```

