---
title: "Multifunctionality-mechanism"
author: "Robert Bagchi"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  github_document:
    toc: yes
    toc_depth: 5
    pandoc_args: --webtex
  html_document:
    keep_md: yes
    theme: readable
    mathjax: default
  pdf_document:
    toc: yes

---

# Rationale

We can think of the relationship between species composition and function provision using the following relationship

$$
\vec{F} =  \vec{N} \cdot \bf{F} 
$$
Where  the rwo vector $\vec{F}$ measures the provision of each of $m$ functions and $\bf{F}$ is a $n \times m$ matrix where each row gives each species' *per capita* contribution to the ecosystem functions. $\vec{N}$ is a row vector of abundances of each species.

The *multifunctionality* part of the model can be examined by expressing each row, $i$ of $\bf{F}$ as

$$
\begin{array}{lll}
\vec{F}_{i.} \sim MVN(\mu, \Sigma) \\
\Rightarrow \bf{F}_{i.} =  \mu + \vec{z} \cdot (\sigma \cdot \Omega^{0.5}\cdot \sigma^t),~ \vec{z} \sim \mathcal{N}(0,1)
\end{array}
$$
where $\Omega$ is the correlation matrix, $\mu$ is the mean of each function, $\sigma$ is the standard deviation of each function. $\vec{z}$ is a row vector of length $m$.

For example, if we assume a unit normal distribution for each function and assume z is an $n \times m$ matrix

$$
\vec{F} =  \vec{N} \cdot \bf{z} \cdot \Omega^{0.5}
$$

We can see how each function depends on the correlation matrix. How that might translate into a directional effect however, I don't yet understand. 

```{r}
n <- 10; m <- 5
Omega <- rethinking::rlkjcorr(1, K =m, eta= 2) ## smaller eta is more correlation among species

z <- matrix(rnorm(n*m), nr=n)

Sigma <-  z %*% diag(m) %*% chol(Omega) %*% t(diag(m))

F_k <- rlnorm(n) %*% Sigma
F_k

```

If we want, it is easy to express this for several communities, perhaps with varying diversity (say 1 - 10 species).

```{r}
library(tidyverse)
sp <- matrix(rlnorm(100), nr=10)
sp[upper.tri(sp)] <- 0
F_k <- sp %*% Sigma

funcs <- data.frame(S= 1:10, f = F_k %*% rep(1/m, m))

ggplot(funcs, aes(x=S, y =f)) + geom_point() + geom_smooth(method="lm") +
  labs(x= "Species richness", y = "mean functioning") + theme_bw()
```


We can the do this for different values of eta.
```{r}
genSpFunc <- function(n, m, eta)
{
  Omega <- rethinking::rlkjcorr(1, K =m, eta= eta) ## smaller eta is more correlation among species
  z <- matrix(rnorm(n*m), nr=n)
  Sigma <-  z %*% diag(m) %*% chol(Omega) %*% t(diag(m))
  return(Sigma)
}
eta <- seq(0.5, 10, len = 5)
dat <- data.frame(eta = rep(eta, each = n), S = rep(1:n, length(eta)),  meanfunc =do.call("c", lapply(eta, function(eta, spA )
  spA %*% genSpFunc(n, m, eta) %*% rep(1/m, m), spA = sp)))
  
ggplot(dat, aes(x = S, y = meanfunc, colour = eta, group = eta)) + geom_point() + 
  geom_smooth(method = "lm") + theme_bw() + scale_colour_viridis_c() + facet_wrap(~eta)

```


It is perhaps also worth folding in the work I did in last year on competition, which would influence $\vec{N}$. We could work out the composition of the community at equilibrium and use that in the equations.

Using Lotka-Volterra theory we can estimate equilibrium abundances of $n$ species given a competition matrix $\mathbf{A}$ as,

$$
N^* = \vec{1} \cdot \mathbf{A}^{-1}
$$
where $\vec{1}$ is a row of 1s with length $n$. 


This means that we can modify the equation for ecosystem function from above by substituting $\vec{N}$ with the equation for $N^*$


$$
\vec{F} =  \vec{1} \cdot \mathbf{A}^{-1} \cdot \bf{F} 
$$

And, if $\bf{z} \sim \mathcal{N}(0, 1)$ is an n x m matrix 

$$
\vec{F} =  \vec{1} \cdot \mathbf{A}^{-1} \cdot ( \mu + \bf{z} \cdot (\sigma \cdot \Omega^{0.5}\cdot \sigma^t))
$$


we have a general formula for the provision of multiple ecosystem functions by multiple competing species. 

Because we probably lose little from assuming functions arise from a unit normal, we could get rid of $\mu$ and $\sigma$ to get a simpler equation 

$$
\vec{F} =  \vec{1} \cdot \mathbf{A}^{-1} \cdot  \bf{z} \cdot \Omega^{0.5}
$$


```{r}

## S = no. species, a_ii is intra-specific comp,, a_ij is interspecific comp
makeCmat  <- function(S, a_ii, a_ij)
{
  Alpha <-   matrix(a_ij, nr=S, nc=S) 
  diag(Alpha) <- a_ii
  return(Alpha)
}
Amat <- makeCmat(n, 0.1, 0.01)

rep(1, n) %*%  solve(Amat) %*% (z)  %*% Omega

```

One thing that emerges from this is that the diversity-multifunctionality relationship is determined by two things, the inverse of the competition matrix and the correlations among species contributions to functions. Indeed, I think we could probably get a pretty good idea of the expected relationships by just looking at how those two matrices interact, but I'm not quite sure how to explore that yet.