---
title: "R Notebook"
output: html_notebook
---

This is script sets up simulations to explore the sensitivity to the MF metric to function distribution, standardization 

```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(GGally)
```

Here I draw a bunch of function values from a range of distributions chosen to cover the range from uniform to strongly skewed so that the function values covers a wide range of evenness values. 

```{r, "funcmat"}
# set seed to replicate analysis:
set.seed(999)

# number of functions
funcnum <- 15

# draw `funcnum` function values from a range of different distributions 
func_lnorm <- 
  sapply(rep(seq(0.01,1,0.01), each = 10), function(x) rlnorm(funcnum, 0, x)) %>% 
  as.matrix 
  
func_beta <- 
  sapply(rep(seq(0.01,1,0.01), each = 10), function(x) rbeta(funcnum, 1, x)) %>% 
  as.matrix 

func_gamma <- 
  sapply(rep(seq(0.01,4,0.01), each = 100), function(x) rgamma(funcnum, x)) %>% 
  as.matrix 

funcmat <- cbind(func_lnorm, func_beta, func_gamma)

```


I standardize all values by their plot maximum after shifting all values to positive in case we unexpectedly drew a negative number. 
I the calculate the *Average*, *evenness factor*, the *effective number of functions* and the *scaled and unscaled multifunctionality*. 

!! Here I standardize each set of funcnum values, normaly one would standardize across set of plots and not each plot would have a maximum of 1. !!

```{r}
#standardize function values
funcmat_st <- apply(funcmat, 2, function(x) {x + min(x)})
funcmat_st <- apply(funcmat, 2, function(x) {x / max(x)})

# calculate MF metrics for each draw of function values
MF_comp <- tibble(
  Average = apply(funcmat_st, 2, mean),
  effN = apply(funcmat_st, 2, vegan::renyi, scales = 1, hill = T)) %>% 
  mutate(even_fac = effN / funcnum) %>% 
  mutate(MF_scaled = even_fac * Average) %>% 
  mutate(MF_effN = effN * Average)

MF_comp %>% 
  ggplot(aes(x = even_fac))+
  geom_histogram()

```


I plot the pairwise comparisions of interest. Note that I make the plots with MF_scaled but I could as well make them with MF_effN. The number of functions doesn't change so MFeffN = MFscaled * funcnum and is linearly related. 
```{r}
p_Av_MFsclaed <- 
ggplot(MF_comp, aes(Average, MF_scaled, colour = even_fac))+
  geom_point(size = 0.5)+
  scale_colour_viridis_c()+
  geom_abline(intercept = c(0,0), slope = 1)+
  coord_fixed()
  
p_evF_MFsclaed <-
ggplot(MF_comp, aes(even_fac, MF_scaled, colour = Average))+
  geom_point(size = 0.5)+
  scale_colour_viridis_c()+
  geom_abline(intercept = c(0,0), slope = 1)+
  coord_fixed()

p_Av_evF <-
ggplot(MF_comp, aes(Average, even_fac, colour = MF_scaled))+
  geom_point(size = 0.5)+
  scale_colour_viridis_c()+
  geom_abline(intercept = c(0,0), slope = 1)+
  coord_fixed()
  
```

arrange and plot plots
```{r}
grid.arrange(p_Av_MFsclaed, p_evF_MFsclaed, p_Av_evF, nrow = 2)
```

# effect of standardizations

```{r}

Funcs <- 
t(funcmat) %>% 
  as.data.frame()
```

```{r, "standardize"}
# extract function names
func.names <- paste("V", 1:funcnum, sep = "")

#choose standardization function
standfunc <- standardizeUnitScale # by max
#standfunc <- standardizeZScore    # Z-score
#standfunc <- standardizeHedges    # Hedges
#standfunc <- standardizeLR        # Log-ratio

# standardize by max
Funcs_max <- apply(Funcs[,func.names], 1, standardizeUnitScale) %>% t()

# Z-score
Funcs_Z <- apply(Funcs[,func.names], 1, standardizeZScore) %>% t()

# Hedges
Funcs_H <- apply(Funcs[,func.names], 1, standardizeHedges) %>% t()

# Log-ratio
Funcs_LR <- apply(Funcs[,func.names], 1, standardizeLR) %>% t()

# alternative standardize between 0 and 1
Funcs_01 <- apply(Funcs[,func.names], 1, function(x) {(x - min(x)) / (max(x) - min(x))}) %>% t()

Func_df_list <- list(Funcs = Funcs,
                     Funcs_max = Funcs_max,
                     #Funcs_Z = Funcs_Z,
                     Funcs_H = Funcs_H,
                     Funcs_LR = Funcs_LR,
                     Funcs_01 = Funcs_01
                     )

effN_list <- 
lapply(Func_df_list, function(x) {
  apply(x, 1, vegan::renyi, scales = 1, hill = 1)
}) 

effN_df <- 
do.call(cbind, effN_list) %>% 
  as.data.frame() 

```

```{r}
p_max <- 
ggplot(effN_df, aes( x = Funcs, y = Funcs_max))+
  geom_point()+
  geom_abline(intercept = c(0,0), slope = 1, col = "red")+
  coord_fixed()

p_H <- 
  ggplot(effN_df, aes( x = Funcs, y = Funcs_H))+
  geom_point()+
  geom_abline(intercept = c(0,0), slope = 1, col = "red")+
  coord_fixed()

p_LR <- 
  ggplot(effN_df, aes( x = Funcs, y = Funcs_LR))+
  geom_point()+
  geom_abline(intercept = c(0,0), slope = 1, col = "red")+
  coord_fixed()

p_01 <- 
  ggplot(effN_df, aes( x = Funcs, y = Funcs_01))+
  geom_point()+
  geom_abline(intercept = c(0,0), slope = 1, col = "red")+
  coord_fixed()

grid.arrange(p_max, p_H, p_LR, p_01)
```

